# Prediction Machine Learning Writeup

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har

##Data Cleaning and Preparation

```{r}
set.seed(1212)
library(gbm)
library(randomForest)
library(caret)
library(ggplot2)
#Data loading
pmlTrain<-read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
pmlTest<-read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))
```

Training data was partitioned and preprocessed using the code described below. In brief, all variables with at least one "NA" were excluded from the analysis. Variables related to time and user information were excluded for a total of 51 variables and 19622 class measurements. Same variables were mainteined in the test data set (Validation dataset) to be used for predicting the 20 test cases provided.

```{r}
## NA exclusion for all available variables
noNApmlTrain<-pmlTrain[, apply(pmlTrain, 2, function(x) !any(is.na(x)))] 
dim(noNApmlTrain)
```

```{r}
## variables with user information, time and undefined
cleanpmlTrain<-noNApmlTrain[,-c(1:8)]
dim(cleanpmlTrain)
```

```{r}
## 20 test cases provided clean info - Validation data set
cleanpmltest<-pmlTest[,names(cleanpmlTrain[,-52])]
dim(cleanpmltest)
```


##Data Partitioning and Prediction Process

The cleaned downloaded data set was subset in order to generate a test set independent from the 20 cases provided set. Partitioning was performed to obtain a 75% training set and a 25% test set.

```{r}
#data cleaning
inTrain<-createDataPartition(y=cleanpmlTrain$classe, p=0.75,list=F)
training<-cleanpmlTrain[inTrain,] 
test<-cleanpmlTrain[-inTrain,] 
#Training and test set dimensions
dim(training)
dim(test)
```


##Results

Random Forest trees were generated for the training dataset using cross-validation. Then the generated algorithm was examnined under the partitioned training set to examine the accuracy and estimated error of prediction. 

###By using 51 predictors for five classes using cross-validation at a 5-fold an accuracy of 99.3% with a 95% CI [0.990-0.995] was achieved accompanied by a Kappa value of 0.9915.

```{r}
fitControl2<-trainControl(method="cv", number=5, allowParallel=T, verbose=T)
rffit<-train(classe~.,data=training, method="rf", trControl=fitControl2, verbose=F)
```

```{r}
predrf<-predict(rffit, newdata=test)
confusionMatrix(predrf, test$classe)
```


```{r}

pred20<-predict(rffit, newdata=cleanpmltest)
# Output for the prediction of the 20 cases provided
pred20
```


###A boosting algorithm (gbm) was also run to confirm and be able to compare predictions. Data is not shown but the boosting approach presented less accuracy (95%) (Data not shown). However, when the predictions for the 20 test cases were compared match was same for both algorimths.

